Machine Learning guided workflow for Ribosome Binding Site engineering

### 2021-05-19 paper/main.tex
change colors, setup subfigure caption, add annotation to tsne

@@

\begin{figure}[!ht]
    \centering
    \includegraphics[scale=0.4]{plots/Main_Paper/swarmplot.pdf}
    \includegraphics[scale=0.4]{plots/Main_Paper/quantplot.pdf}
    \includegraphics[scale=0.4]{plots/Main_Paper/histogram.pdf}
    \includegraphics[scale=0.4]{plots/Main_Paper/tsneplot.pdf}
    \caption{\cheng{May need to use subfigure to label the figures as A,B,C,D.}
    \textbf{TIRs of RBS groups examined in this study.} 
    The TIR results in all subplots are shown normalised to the respective benchmark sequence sample which acts as internal standard, that is TIR of a given RBS is divided by TIR of the benchmark RBS run in the same plate. 
    \textbf{A)} Swarm plot showing the obtained TIRs divided into RBS groups.
    BPS-NC: base-by-base changes in the non-core region. 
    BPS-C: base-by-base changes in the core region. 
    UNI: Randomly generated sequences with uniform distribution. 
    PPM: Randomly generated sequences with distribution following the PPM for all natural RBS in \emph{E. coli}. 
    Bandit-0/1/2/3 - Bandit algorithm generated results for Round 0, 1, 2 and 3 respectively.
    SD - Shine-Dalgarno sequence.
    Dash line is set to 1 and represents the averaged benchmark sequence TIR for that group. 
    BN - benchmark sequences for all plates. 
    They are not all exactly 1 due to them being shown as separate samples rather than averages.
    \textbf{B)} Line plot showing TIR obtained in a given quantile of results divided into groups as in A).
    % save the random groups which were shown together due to similar distributions.
    UNI and PPM are merged into Random group and BPS-NC was removed due to changes being made outside the core in this group. 
        \textbf{C)} Histogram with kernel density estimations (KDE) showing distributions of TIRs for Bandit groups.
    \textbf{D)} tSNE plot showing the relative distances between sequences in our design spaces as calculated by our kernel function (weighted degree kernel with shift). 
    The area size of the circle represents the experimentally obtained TIR for measured groups.}
    \label{fig: Swarmplot and Quantplot}
\end{figure}

@@

\begin{figure}
     \centering
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \includegraphics[scale=1.2]{plots/Main_Paper/TOP30_logo.pdf}
         \caption{}
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \includegraphics[scale=0.5]{plots/Main_Paper/Hd_Heatmap.png}
         \caption{}
     \end{subfigure}
     \caption{\textbf{Analysis of RBS positional information} A) Sequence logo calculated for the Top 30 tested sequences. B) Heatmap showing what edit (Hamming) distance is required for positive change in TIR for RBS with high and medium TIR. The temperature scale shows the difference between a given RBS on y axis and the RBS with strongest TIR at the given distance.}
     \label{fig:Library characteristics}
\end{figure}

>>>>

% \captionsetup[subfigure]{position=top, labelfont=bf,textfont=normalfont,singlelinecheck=off,justification=raggedright}
\captionsetup[subfigure]{font={bf,small}, skip=1pt, singlelinecheck=false}
\renewcommand{\thesubfigure}{\Alph{subfigure}}

@@

\begin{figure}[!ht]
    \centering
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \caption{}
        \includegraphics[scale=0.3]{plots/Main_Paper/swarmplot.pdf}
    \end{subfigure}
    % \hfill
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \caption{}
        \includegraphics[scale=0.3]{plots/Main_Paper/quantplot.pdf}
    \end{subfigure}
    % \hfill
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \caption{}
        \includegraphics[scale=0.3]{plots/Main_Paper/histogram.pdf}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \caption{}
        \includegraphics[scale=0.3]{plots/Main_Paper/tsneplot.pdf}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \caption{}
        \includegraphics[scale=0.3]{plots/Main_Paper/swarmplot_proj.pdf}
    \end{subfigure}
    % \includegraphics[scale=0.3]{plots/Main_Paper/swarmplot.pdf}
    % \includegraphics[scale=0.3]{plots/Main_Paper/quantplot.pdf}
    % \includegraphics[scale=0.3]{plots/Main_Paper/swarmplot_proj.pdf}
    % \includegraphics[scale=0.3]{plots/Main_Paper/histogram.pdf}
    % \includegraphics[scale=0.4]{plots/Main_Paper/tsneplot.pdf}
    \caption{
    \textbf{TIRs of RBS groups examined in this study.} 
    The TIR results in all subplots are shown normalised to the respective benchmark sequence sample which acts as internal standard, that is TIR of a given RBS is divided by TIR of the benchmark RBS run in the same plate. 
    \textbf{A)} Swarm plot showing the obtained TIRs divided into RBS groups.
    BPS-NC: base-by-base changes in the non-core region. 
    BPS-C: base-by-base changes in the core region. 
    UNI: Randomly generated sequences with uniform distribution. 
    PPM: Randomly generated sequences with distribution following the PPM for all natural RBS in \emph{E. coli}. 
    Bandit-0/1/2/3 - Bandit algorithm generated results for Round 0, 1, 2 and 3 respectively.
    SD - Shine-Dalgarno sequence.
    Dash line is set to 1 and represents the averaged benchmark sequence TIR for that group. 
    BN - benchmark sequences for all plates. 
    They are not all exactly 1 due to them being shown as separate samples rather than averages.
    \textbf{B)} Line plot showing TIR obtained in a given quantile of results divided into groups as in A).
    % save the random groups which were shown together due to similar distributions.
    UNI and PPM are merged into Random group and BPS-NC was removed due to changes being made outside the core in this group. 
        \textbf{C)} Histogram with kernel density estimations (KDE) showing distributions of TIRs for Bandit groups.
    \textbf{D)} tSNE plot showing the relative distances between sequences in our design spaces as calculated by our kernel function (weighted degree kernel with shift). 
    The area size of the circle represents the experimentally obtained TIR for measured groups.
    \textbf{E)} Exploitation v.s. Exploration for Bandit 1-3. Colors approaching to blue represents exploitation, otherwise represents exploration.}
    \label{fig: Swarmplot and Quantplot}
\end{figure}

@@

\begin{figure}
     \centering
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \caption{}
         \includegraphics[scale=1.2]{plots/Main_Paper/TOP30_logo.pdf}
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \caption{}
         \includegraphics[scale=0.5]{plots/Main_Paper/Hd_Heatmap.png}
     \end{subfigure}
     \caption{\textbf{Analysis of RBS positional information} A) Sequence logo calculated for the Top 30 tested sequences. B) Heatmap showing what edit (Hamming) distance is required for positive change in TIR for RBS with high and medium TIR. The temperature scale shows the difference between a given RBS on y axis and the RBS with strongest TIR at the given distance.}
     \label{fig:Library characteristics}
\end{figure}

### 2021-05-19 paper/methods.tex
change colors, setup subfigure caption, add annotation to tsne

Since experimentally labelling sequences is time-consuming, it is unrealistic to recommend sequence sequentially (i.e. one-by-one) and then wait for the label to be tested and used to improve the model.
Instead, we can recommend RBS sequences in a batch of size $m$. 
One naive approach is to 
%train GP on sequences with known TIR labels and
recommend sequences in design space with top $m$ UCB scores, as shown in Figure \ref{fig:batch rec}(a) ($m = 2$).
However, this approach may end up recommending similar sequences in the same local maximum (e.g. $x = 2, x =2.5$ in this example). 
However, since GPR assumes similar sequences would have similar labels (e.g. by knowing $x=2$ we can gain information of $x=2.5$ as well), we prefer to not waste time and money on labelling sequences with high similarities in the same batch.

@@

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \includegraphics[scale=.4]{plots/Recommend_2_Data_Point_Top_2_UCB.pdf}
    \end{subfigure}
    \vskip\baselineskip
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \includegraphics[scale=0.4]{plots/GP-BUCB_Recommend_1st_Data_Point.pdf}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \includegraphics[scale=0.4]{plots/GP-BUCB_Recommend_2nd_Data_Point.pdf}
    \end{subfigure}
    \caption{Batch Recommendation Illustration. We use the batch size of 2, with 5 initial observations. The design space is 24 uniformly distributed points in the range [-2,10], i.e. {-2, -1.5, -1, ..., 9.5, 10}.
    The confidence interval are shown with predicted mean of $\pm$ 1.96 standard deviation.
    (a) Top UCB recommendations. The recommendations are 2 data points with top UCB scores, constructed with GP predictions. 
    (b)(c) Batch UCB recommendations. (b) shows the first recommended sequence, (c) shows the new predicted confidence interval and the second recommendation based on that.}
    \label{fig:batch rec}
\end{figure}

A key property of Gaussian Process regression is that the predictive standard deviation depends only on observed points (i.e. features), but not on the labels of those observed points. 
% We illustrate this fact by comparing the prediction with the true function value of the new recommended point (Figure \ref{fig: Recommend 1 Data Point (New Observation)}) and the prediction with the predicted value of the new recommended point (Figure \ref{fig: Recommend 1 Data Point (Predicted Mean)}).
% As we can see, the predicted confidence interval is the same for the two plots. 
One can make use of this property to design batch upper confidence bound (BUCB) algorithm \cite{desautels2012parallelizing}.
The strategy is to recommend sequences sequentially by updating the UCB score with the updated predicted standard deviation with the previously recommended points.
% Figure \ref{fig: Recommend 2 Data Points: GP-BUCB.} shows an example of batch UCB recommendation. 
% The plot shows the predicted mean and standard deviation after observing 4 data points. 
As illustrated in Figure \ref{fig:batch rec} (b), the algorithm recommends the data point with maximum UCB score based on the predictions over initial 5 observations.  
Then we add the recommended data point ($x = 2$) into the training dataset with the predicted mean of that point as label (note it is not the true label, i.e. observation), and update the predicted standard deviation and then we finally update the UCB scores. 
The second data point is then recommended based on the new UCB scores.   
As we can see in Figure \ref{fig:batch rec}(c), since we assume we have observed $x = 2$, then the new predicted standard deviation of the data points in design space around $x =2$  decreases, so instead of recommending a similar data point $x = 2.5$, we recommend the data point which is in a different maximum area with potentially high labels ($x = 8$).     
As summary, the exploration efficiency is improved since the recommended sequences in one batch will tend to be in different design area so that the information gain is maximised.

>>>>

Since experimentally labelling sequences is time-consuming, it is unrealistic to recommend sequence sequentially (i.e. one-by-one) and then wait for the label to be tested and used to improve the model.
Instead, we can recommend RBS sequences in a batch of size $m$. 
One naive approach is to 
%train GP on sequences with known TIR labels and
recommend sequences in design space with top $m$ UCB scores, as shown in Figure \ref{fig:batch rec}(A) ($m = 2$).
However, this approach may end up recommending similar sequences in the same local maximum (e.g. $x = 2, x =2.5$ in this example). 
However, since GPR assumes similar sequences would have similar labels (e.g. by knowing $x=2$ we can gain information of $x=2.5$ as well), we prefer to not waste time and money on labelling sequences with high similarities in the same batch.

@@

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \caption{}
        \includegraphics[scale=.4]{plots/Recommend_2_Data_Point_Top_2_UCB.pdf}
    \end{subfigure}
    \vskip\baselineskip
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \caption{}
        \includegraphics[scale=0.4]{plots/GP-BUCB_Recommend_1st_Data_Point.pdf}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \caption{}
        \includegraphics[scale=0.4]{plots/GP-BUCB_Recommend_2nd_Data_Point.pdf}
    \end{subfigure}
    \caption{Batch Recommendation Illustration. We use the batch size of 2, with 5 initial observations. The design space is 24 uniformly distributed points in the range [-2,10], i.e. {-2, -1.5, -1, ..., 9.5, 10}.
    The confidence interval are shown with predicted mean of $\pm$ 1.96 standard deviation.
    (A) Top UCB recommendations. The recommendations are 2 data points with top UCB scores, constructed with GP predictions. 
    (B)(C) Batch UCB recommendations. (B) shows the first recommended sequence, (C) shows the new predicted confidence interval and the second recommendation based on that.}
    \label{fig:batch rec}
\end{figure}

A key property of Gaussian Process regression is that the predictive standard deviation depends only on observed points (i.e. features), but not on the labels of those observed points. 
% We illustrate this fact by comparing the prediction with the true function value of the new recommended point (Figure \ref{fig: Recommend 1 Data Point (New Observation)}) and the prediction with the predicted value of the new recommended point (Figure \ref{fig: Recommend 1 Data Point (Predicted Mean)}).
% As we can see, the predicted confidence interval is the same for the two plots. 
One can make use of this property to design batch upper confidence bound (BUCB) algorithm \cite{desautels2012parallelizing}.
The strategy is to recommend sequences sequentially by updating the UCB score with the updated predicted standard deviation with the previously recommended points.
% Figure \ref{fig: Recommend 2 Data Points: GP-BUCB.} shows an example of batch UCB recommendation. 
% The plot shows the predicted mean and standard deviation after observing 4 data points. 
As illustrated in Figure \ref{fig:batch rec} (B), the algorithm recommends the data point with maximum UCB score based on the predictions over initial 5 observations.  
Then we add the recommended data point ($x = 2$) into the training dataset with the predicted mean of that point as label (note it is not the true label, i.e. observation), and update the predicted standard deviation and then we finally update the UCB scores. 
The second data point is then recommended based on the new UCB scores.   
As we can see in Figure \ref{fig:batch rec}(C), since we assume we have observed $x = 2$, then the new predicted standard deviation of the data points in design space around $x =2$  decreases, so instead of recommending a similar data point $x = 2.5$, we recommend the data point which is in a different maximum area with potentially high labels ($x = 8$).     
As summary, the exploration efficiency is improved since the recommended sequences in one batch will tend to be in different design area so that the information gain is maximised.
